---
title: "Data Viz"
author: "Felicia Cruz"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use
library(tidyverse)
library(here)
library(janitor)

# Specifically for data viz 
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

```

Histogram (really bad) of lizard weights 
```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange", 
                 color = "black",
                 size = 1)

```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             color = "blue",
             fill = "cyan",
             size = 4,
             alpha = 0.4)
```
Bubble plot where the color of the points changes based on common name and the size of points varies based on total_length
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, 
                 size = total_length),
             shape = 22,
             fill = "black") +
  theme_light()


```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color= common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 4)

```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color= common_name)) +
  facet_grid(sex ~ tail)

```
Find total lizard counts by common name
```{r}
lizard_counts <- lizards %>%
  group_by(common_name) %>% 
  summarize(count = n())
  
lizard_counts <- lizards %>% 
  count(common_name)

lizards_counts_cn_tail <- lizards %>% 
  count(common_name, tail)
```


```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) + 
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "lizard counts") +
  theme_light()

```
Try converting common name to an ordered factor outside of ggplot.
```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, n))
```

## Axis scale breaks, limits, and labels 

Scatterplot of total_length vs weight 
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 50, 500),
                     limits = c(0, 500),
                     expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))

```
Trasnform date column to class Date, then find counts on observations by date 
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)
```

```{r}
ggplot(data = lizard_counts, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "'%y")


```


```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("green", "blue", "purple"), 
                     breaks = c(0, 10, 20, 40, 50) )

```

### Update a color scheme using a palette in paletteer 
Make a horizontal box plot
```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(lizards_fct, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "miscpalettes::sunset")

```

```{r}
ggplot(lizards_fct, aes(x = total_length, y = weight)) +
  geom_point()+
  theme(
   panel.grid = element_blank(),
   plot.background = element_rect(fill = "cyan"),
   axis.text.x = element_text(color = "orange"),
   axis.title = element_text(color = "yellow", size = 15)
  )

```





